---
title: "ECON0128 Assignment 2"
author: "Fayyaz Baig"
date: "2024-10-23"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
# Clear Environments and Load necessary libraries
rm(list=ls())
library(tidyverse)
library(texreg)
library(DBI)
library(RPostgres)

```


## Frisch Waugh Lovell Theorem

Data Loading
```{r}
#We load the dataset from Dropbox.
# Load dataset from Dropbox
county_data <- read.csv(paste0("https://www.dropbox.com/scl/fi/lumi2zlvbfcd2qsfnomqh/", 
"county_data_24.csv?rlkey=0jin81pffu9pohq6vquudbi1b&e=1&dl=1"))

# Check if dataset has loaded correctly
str(county_data)
head(county_data)
names(county_data)

```

## Data Processing

We transform the variable y_hh_med into its logarithmic form.

```{r}
county_data <- county_data %>% 
  mutate(log_y_hh_med = log(y_hh_med))

```


Multiple Regression
We perform a multiple regression of rep_share on log_y_hh_med and white_share.
```{r}
# Perform a multiple regression of rep_share on log_y_hh_med and white_share
model_lm <- lm(rep_share ~ log_y_hh_med + white_share, data = county_data)

# Summary of the model
screenreg(model_lm)

```

## Application of Frisch Waugh Lovell (FWL) Theorem
We demonstrate the FWL theorem by conducting residual regressions.

```{r}
# Regress rep_share on white_share 
model_fwl1 <- lm(rep_share ~ white_share, data = county_data)
res1 <- model_fwl1$residuals

# Regress log_y_hh_med on white_share
model_fwl2 <- lm(log_y_hh_med ~ white_share, data = county_data)
res2 <- model_fwl2$residuals

# Regress the residuals
model_fwl3 <- lm(res1 ~ res2)

# Display results of both models
screenreg(list(model_lm, model_fwl3))

```

## Visualizing the FWL Theorem
We create a scatter plot of the residuals from the two regressions using ggplot2.

```{r}
# Create a scatter plot of the residuals
ggplot(aes(x = res2, y = res1), data = county_data) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Plot of Residuals",
         x = "Residuals Model 2",
         y = "Residuals Model 1") +
    theme_minimal()

```

The plot demonstrates the FWL theorem by showing the relationship between the two residuals. The slope should be equal to the coefficient of the log_y_hh_med variable in the original model.

## Direct Regression of rep_share on res2

We regress rep_share on res2 to further confirm the FWL theorem.

```{r}
# Regress rep_share directly with res2
model_fwl4 <- lm(rep_share ~ res2, data = county_data)

# Display comparison of models
screenreg(list(model_lm, model_fwl3, model_fwl4))


```

## Task 2: Database Interaction

Part 1: Setting Up and Connecting to Database
First, we install and load the necessary packages to connect to a PostgreSQL database.

```{r}
# Connect to the Postgres database
data <- dbConnect(
  RPostgres::Postgres(),
  dbname = "movies_database",
  host = "34.89.96.247",
  port = "5432",
  user = "postgres",
  password = "UCL_ECON_0128"
)

# Check if the connection has been established
dbListTables(data)
dbListFields(data, "movies")


```

## Part 2: Querying the Database

Method 1: SQL Queries

We retrieve all records from the ratings table where the rating is 5.

```{r}
# Get all records from the ratings table where the rating is 5
ratings_query <- dbSendQuery(data, "SELECT * FROM ratings WHERE rating = 5")
ratings_results <- dbFetch(ratings_query)
dbClearResult(ratings_query)

# Display results as a tibble
ratings_results <- as_tibble(ratings_results)

```

Method 2: Queries with dbplyr

We filter and arrange the ratings table using dbplyr.

```{r}
# Filter the ratings table for records where the rating is 5 and the timestamp is after January 1, 2016
ratings <- tbl(data, "ratings")

ratings_filtered <- ratings %>%
  filter(rating == 5, timestamp > "2016-01-01") %>%
  arrange(timestamp) %>%
  collect()
```


Part 3: Working with Multiple Tables
We rank the top 10 movies according to the number of ratings 4.5 and above in 2007.

```{r}
# Create a table with the number of ratings 4.5 and above for each movie in 2007
#movies <- tbl(data, "movies")
#ratings_movies_merged <- ratings %>%
#  group_by(movie_id) %>%
#  summarise(average_rating = mean(rating, na.rm = TRUE), n = as.numeric(n())) %>%
#  left_join(movies, by = "movie_id")

# Filter and arrange the data
#top_movies_2007 <- ratings_movies_merged %>%
#  filter(average_rating >= 4.5, year(timestamp) == 2007) %>%
#  arrange(desc(n)) %>%
#  select(movie_id, title, n, average_rating) %>%
#  head(10)

# Display top 10 movies
#top_movies_2007
```
